---
title: "Lab Notebook"
author: "Beniamino Green"
output:
  pdf_document:
    extra_dependencies: ["float"]
link-citations: yes
bibliography: bib.bib
---

```{r, echo=F, include=F}
library(tidyverse)
library(stargazer)
library(dotwhisker)
library(arrow)
library(corrplot)
load("data/models.Rda")
data <- read_parquet("data/cleaned_data.parquet")
```

# Experiment Summary:

This notebook contains data analysis performed on the "social polarization"
conjoint experiment. As a refresher, respondents were shown two profiles, which
had the following attributes:

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|}
        \hline
        Person A & Person B \\
        \hline
        Party & Party \\
        Ideology & Ideology \\
        Race / Ethnicity & Race / Ethnicity \\
        Regional Background & Regional Background   \\
        Education & Education \\
        Income & Income \\
        Gender & Gender \\
        Religion & Religion \\
        \hline
    \end{tabular}
\end{table}

Where each attribute was randomized and could take one take on one the
following values:

* Party: A number, 1-3 (1=Democratic, 2=Independent, 3=Republican)
* Ideology: A number, 1-3 (1=Liberal, 2=Moderate, 3=Conservative)
* Race/Ethnicity: either White, Black, Hispanic, or Asian.
* Regional Background: Northeast,Midwest, South, or West
* Education: One of HS grad, 2-year college, 4-year college, or Post-grad. Coded as a numeric vector 0-3, where 0 is HS grad, and $ is post-grad.
* Income: A number, 1-4 (Under $40K=0, $40-80K=2, $80-120K=3, Over $120K=4)
* Gender: Male (0) or Female (1)
* Religion: Christian, Jewish, Muslim, or Atheist

Respondents were then asked:

\begin{quote}
    On a scale or 1 (lowest) to 7 (highest), how favorably do you view each of the
    individuals profiled?

    Indicate which individual (A or B) you would prefer in each of the following scenarios:
    \begin{itemize}
        \item Having as a friend
        \item Having as a neighbor
        \item Having as a son/daughter-in-law
    \end{itemize}
\end{quote}

The first question (On a scale of 1 to 7 \dots) yields a ratio-level measure of
favorability, which will be referred to in this document as the ratio-measure.
The other two questions forced respondents to pick one of the two profiles that
they would prefer to have as a certain kind of acquaintance, yielding
binary-choice data.

# Exploratory Data Analysis:

In this section, I perform some exploratory data analysis to better understand
the data.

## How Similar Are Answers to the Three Conjoint Questions?

First, I want to understand to what degree are the answers to the three
conjoint questions correlated, and do the questions tap into different
underlying dimensions?  One hypothesis might be that respondents are answering
each of the three questions identically, based on which person they like the
most in a general sense.

To understand this, I first create a correlation plot showing correlations
between answers to the three questions, which can be seen below:

```{r, echo=F}
conjoint_outcomes <- data %>%
    select(starts_with("Q2"))

# conjoint_outcomes <- as.data.frame(conjoint_outcomes == 2)
# names(conjoint_outcomes) <- c("Friend", "Neighbor", "S/DIL")

# corrplot.mixed(cor(conjoint_outcomes))

# glm(Friend ~ Neighbor, family=binomial(link="logit"), data=conjoint_outcomes)
# glm(`S/DIL` ~ Neighbor, family=binomial(link="logit"), data=conjoint_outcomes)

# n_true <- apply(conjoint_outcomes,1,sum)
# mean(n_true==0 | n_true ==3)
```

The correlations range between .54 and .67 indicating relatively strong
positive correlations between answers to the three questions.

Another way to understand how closely related the three questions are is to
look at the percentage of people that preferred one profile to another on all
three conjoint questions - 67%. This suggests that the questions do measure
similar concepts, but that they do not measure exactly the same thing, as
respondents did not uniformly prefer one profile over another in 33% of
conjoints.

# Data Analysis

In this section, I proceed to the substantive data analysis. I first regress
the interval-level measure of likability on the attributes of the conjoint
profile. As a note, the reference category for the region variable is
"Northeast," the reference category for the religion variable is "Christian,"
and the reference category for race is "White."  Party is measured on a 1-3
scale, where 1 is a Democrat, 2 is an independent, and 3 is a Republican.

The coefficients associated with a profile being female or black are positive
and significant, with coefficient estimates of .09 each. This indicates that a
profile who is female or black will receive a predicted rating .09 points
higher than an identical profile who is not Black or not Female. As the
attribute profiles have been randomized, these estimates are unbiased in
expectation, and can be interpreted causally.

The coefficients associated with being Muslim and Atheist are both significant
and negative, indicating that relative to Christians, Muslims and Atheists are
rated less-favorably by respondents.

```{r, echo=F, results="asis"}

stargazer(
          ratio_model_2, header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Liberal",
                               "Conservative",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
          dep.var.labels = "Favorability Rating (1-7)",
          title = "Regression of Favorability on Profile Attributes",
          table.placement="H"
)
```

I display the coefficients and standard errors from this model in a coefficient
plot, shown below. The plot contains no new information, but is easier to take
in at-a-glance.

```{r, echo=F}

dwplot(ratio_model_2) +
    ggtitle("Coefficient Plots For Model 1") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Liberal",
                               "Conservative",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    )
```

## Analysis of Binary-Choice Data

Secondly, I run regress the outcomes from the binary choice questions on
conjoint profile characteristics. Coefficients in these regressions can be
interpreted as average marginal component effects, or the difference in
expected vote share^[taken over the joint distribution of candidate attributes]
between two candidates that have a one-unit difference in the predictor of
interest.

As a note, the AMCE is not equal to the expected difference in vote shares in
an "election" between the two candidates that differ by one unit on a
predictor. Rather, it is equal to the difference in the average vote share each
candidate would receive if they were to compete against all other candidates in
the population.

To give an example, the coefficient associated with a profile being black in
the first regression model suggests that a black profile will receive on
average 3.5% more votes than an otherwise-identical white profile, across all
elections.


```{r, echo=F, results="asis"}

stargazer(
          acme_model_1, acme_model_2, acme_model_3,
          header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
      dep.var.labels = c("Friend","Neighbor", "S/DIL")
)

```

I also present the point estimates and confidence intervals in a graphical
form, as before. Again, the graph below contains no new information, but is
slightly easier to read at-a-glance.

```{r, echo=F}

dwplot(
          list("Friend Model"= acme_model_1,
               "Neighbor Model" = acme_model_2,
               "S/DIL Model" = acme_model_3)
       ) +
    ggtitle("Coefficient Plots For Model ACME models") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    ) +
    scale_x_continuous(limits=c(-.2,.2))

```

# Split-Sample Estimates:

In this section, I report the results of split-sample estimates of the effects
of profile attributes on the 1-7 likability scale question. In each section, I
repeat the same analyses as above on subgroups of the sample (Republicans,
Democrats, etc).

It is important to remember in reading the regression tables that The
difference between a statistically significant and a statistically
insignificant point estimate may not itself be statistically significant^[see [here](https://statmodeling.stat.columbia.edu/2005/06/14/the_difference/) for more].

This is to say that if a coefficient is significant for one group but
insignificant for another, this does not mean that the difference in the
effects is significantly different between the two groups. To calculate the
standard error associated with the difference in the two groups, we must use
the following formula:

$${S.E.}_\text{Difference} = \sqrt{{S.E.}_\text{Group 1}^2 + {S.E.}_\text{Group 2}^2}$$


## Democrats VS Republicans:

First, I regress likability measured on the 1-7 scale on the conjoint profile
attributes among just Democratic-leaning respondents, and just
Republican-leaning respondents.

The coefficients for party and ideology are significantly different, and in the
directions one would expect; Republican respondents like republican profiles
more than democratic ones, while Democrats like republican profiles less than
otherwise-identical democratic ones.

The difference between the coefficient for a Black profile is positive for
democrats and negative for republicans and the difference is statistically
significant at the $p=.05$ level, suggesting that democrats view Black profiles
more favorably than republican respondents.

Finally, the difference between how Democrats and Republicans view Muslim and
Atheist (although not Jewish) profiles are also statistically significant.
Republicans view Jewish, Muslim and Atheist profiles significantly more
negatively than otherwise-identical Christian profiles, and the differences
between the estimates for Democrats and Republicans are statistically
significant for the Muslim and Atheist profiles.

```{r, echo=F, results="asis"}
stargazer(
          dem_ratio_model_1, indep_ratio_model_1, rep_ratio_model_1,
          header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
          title = "Regression of Favorability on Profile Attributes",
          dep.var.labels = "Favorability Rating (1-7)",
          column.labels = c("Democrats", "Independents","Republicans"),
          table.placement="H"
)
```

```{r, echo=F}
dwplot(
          list(
               "Democrats"= dem_ratio_model_1,
               "Independents"= indep_ratio_model_1,
               "Republicans" = rep_ratio_model_1)
       ) +
    ggtitle("Coefficient Estimates For Republicans / Democrats") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    )
```

 # Results by Ideology

```{r, echo=F, results="asis"}
stargazer(
          liberal_ratio_model_1, moderate_ratio_model_1, conservative_ratio_model_1,
          header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Liberal",
                               "Conservative",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
          title = "Regression of Favorability on Profile Attributes",
          dep.var.labels = "Favorability Rating (1-7)",
          column.labels = c("Liberals", "Moderates","Conservatives"),
          table.placement="H"
)
```

```{r, echo=F}
dwplot(
          list(
               "Liberals"= liberal_ratio_model_1,
               "Moderates"= moderate_ratio_model_1,
               "Conservatives" = conservative_ratio_model_1)
       ) +
    ggtitle("Coefficient Estimates For Republicans / Democrats") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Liberal",
                               "Conservative",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    )
```

## Men VS Women

I repeat the same split-sample analysis for Men vs Women. The interpretations
of the coefficients are the same as above.

Men and Women appear to react to the partisanship of a profile in different
ways; while both appear to react negatively to republican profiles on average,
Women appear to react more negatively, and the difference is statistically
significant.

```{r, echo=F, results="asis"}
stargazer(
          male_ratio_model_1, female_ratio_model_1,
          header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
          title = "Regression of Favorability on Profile Attributes",
          dep.var.labels = "Favorability Rating (1-7)",
          column.labels = c("Men", "Women"),
          table.placement="H"
)
```

```{r, echo=F}
dwplot(
          list("Men"= male_ratio_model_1,
               "Women" = female_ratio_model_1)
       ) +
    ggtitle("Coefficient Estimates For Men / Women") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    )
```

## White VS Non-White Respondents

Finally, I run the same regressions on white vs non-white respondents. The
differences between the two groups are not statistically significant.

```{r, echo=F, results="asis"}
stargazer(
          white_ratio_model_1, non_white_ratio_model_1,
          header=F,
          covariate.labels = c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist",
                               "Intercept"
                               ),
          title = "Regression of Favorability on Profile Attributes",
          dep.var.labels = "Favorability Rating (1-7)",
          column.labels = c("White Respondents", "Non-White Respondents"),
          table.placement="H"
)
```

```{r, echo=F}
dwplot(
          list("White Respondents"= white_ratio_model_1,
               "Non-White Respondents" = non_white_ratio_model_1)
       ) +
    ggtitle("Coefficient Estimates for White / Non-White Respondents") +
    geom_vline(xintercept=0, linetype=2, alpha=.6) +
    xlab("Coefficient Estimate") +
    ylab("Variable") +
    scale_y_discrete(
                     labels=rev(c(
                               "Democrat",
                               "Republican",
                               "Ideology",
                               "Race - Black",
                               "Race - Asian",
                               "Ethnicity - Hispanic",
                               "Region - Midwest",
                               "Region - South",
                               "Region - West",
                               "Education",
                               "Income",
                               "Female",
                               "Religion - Jewish",
                               "Religion - Muslim",
                               "Religion - Atheist"
                              ))
    )
```


## Created Hierarchical Models from @Jensen_2021

Finally, I have written hierarchical Bayesian models to understand
heterogeneity in effects across the popultaion. I have coded and tested the
models, but not yet applied them to the survey data.

All the models take the same form as before, modelling responses as a linear
combination of profile attributes. However, in these models, the effect of each
profile attribute also modelled as a linear combination of the demographic
characteristics of each respondent.

Specifically, I model the rating that a profile, recieves, $Y$ from a person, $i$ as:

$$Y = \beta_i^T X$$

Where $\beta_i$ is a vector that describes the effect each attribute of a
conjoint profile has on a specific respondent's rating. Each $\beta_{ij} \in
\beta_i$ comes from a linear combitation of respondent attributes and
population weights.

$$\beta_{i} = \gamma^T Z_j$$


# References

